Measurement of a dwarf galaxy rotational velocity from its Lyman-alpha line emission

### 2016-06-03 paper/letter.tex
new abstract

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Þú ert jörðin}
\\
\\
JEFR$^{1}$, MCRG$^1$, JNGC$^2$, MG$^3$, MD$^3$
\\
\\
\scriptsize
{$^1$ Bogota
\\
$^2$ Tucson
\\
$^3$ Oslo
\normalsize
\\
\\
\textbf{
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe.    
Before the observational detection of the first galaxies becomes
reality, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
features in its Lyman-$\alpha$ emission that had evaded theoretical
interpretation so far. 
We show that these special features, a symmetric triple peaked
emission line, can be explained by gas bulk rotation.  
We find that the Lyman-$\alpha$ emission region in \tol\ should
have a rotational velocity of $V_{r}=300$ km s$^{-1}$ and a neutral
Hydrogen column density of $\log N_{HI} / \mathrm{atoms\ cm}^{-2} =
  20.5$.   
Using archival observational information about \tol\ we find that the
diameter for that emission region diameter should be in the range of
$110$ pc $<D<$ $340$ pc and its total dynamical mass   should be
between $2.1\times 10^{9}$M$_{\odot}$ $<M_D<$  $6.6\times
10^{9}$M$_{\odot}$.  
This dynamical mass is at least $16\pm 9$ times
larger than the neutral mass hydrogen.
We argue that a possibility to explain the excess in
dynamical mass is the presence of a super-massive black hole. }

>>>>

\begin{document}
\pagestyle{empty}
\noindent
\textbf{Þú ert jörðin}
\\
\\
JEFR$^{1}$, MCRG$^1$, JNGC$^2$, MG$^3$, MD$^3$
\\
\\
\scriptsize
{$^1$ Bogota
\\
$^2$ Tucson
\\
$^3$ Oslo
\normalsize
\\
\\
\textbf{
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe.    
Before the observational detection of the first galaxies becomes
reality, CDGs are the best systems to test our ideas on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs, \tol, which presents
features in its Lyman-$\alpha$ emission that had evaded theoretical
interpretation so far. 
We show that these special features, a symmetric triple peaked
emission line, could be explained by two different models: an
homogeneous sphere undergoing gas bulk rotation or a clumpy outflowing
multiphase inter-stellar medium.
Using archival observational information about \tol\ we 
find that both models require a total dynamical mass in the range
$2.5\times 10^{9}$M$_{\odot}$ $<M_D<$  $4.5\times 10^{9}$M$_{\odot}$.  
This dynamical mass is at least $15\pm 5$ times
larger than the neutral mass hydrogen.
We argue that a possibility to explain the excess in
dynamical mass is the presence of a super-massive black hole.}
